---
title: "Project"
author: "Kevin Choi"
date: '2022-11-13'
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Matrix)
library(tidyverse)
library(here)
library(haven)
library(lme4)
library(lmerTest)  
library(psych)
library(ipumsr)
library(dplyr)
library(modelsummary)
library(apaTables)

```


## Data Cleaning

```{r data}
data <- read_sav(here("data_files", "pj_data.sav"))


data<- data %>% #don't need household data 
   filter(RECTYPE != 1)

data_filled<- data %>% 
  group_by(CASEID) %>%
  mutate(FULLPART = FULLPART[1],
         DAY=DAY[1],
         AGE=AGE[1],
         SEX=SEX[1],
         MARST=MARST[1],
         UHRSWORKT=UHRSWORKT[1],
         EARNWEEK=EARNWEEK[1],
         PAIDHOUR=PAIDHOUR[1],
         HH_NUMOWNKIDS=HH_NUMOWNKIDS[1],
         WB_RESP=WB_RESP[1],
         Leisure_All=Leisure_All[1])%>%
  subset(select = -c(PERNUM, LINENO, WT06, IND2_CPS8, YEAR, AWBWT)) %>% 
  filter(RECTYPE!=2) #after pasting values, remove rows with only level 2
head(data_filled)

#Recoding Hourly Status Variable (1=Not Paid Hourly, 2 = Paid Hourly)
data_filled$PAIDHOUR <- as.numeric(data_filled$PAIDHOUR)
data_filled <- data_filled %>%
  mutate(PAIDHOUR=recode(PAIDHOUR, 
                    '1' = 2,
                    '2' = 1))

#Recoding Day of Week (1=Weekday, 2=Weekend)
data_filled$DAY <- as.numeric(data_filled$DAY)
data_filled <- data_filled %>%
  mutate(DAY=recode(DAY, 
                    '1' = 2,
                    '2' = 1,
                    '3' = 1,
                    '4' = 1,
                    '5' = 1,
                    '6' = 1,
                    '7' = 2))

#Recoding Marital Stauts (1=Not Married, 2= Married)
data_filled$MARST <- as.numeric(data_filled$MARST)
data_filled <- data_filled %>%
  mutate(MARST=recode(MARST, 
                    '1' = 2,
                    '2' = 2,
                    .default = 1))

#Recoding Well-Being Measures
data_filled$MEANING <- as.numeric(data_filled$MEANING)
data_filled <- data_filled %>%
  mutate(MEANING=recode(MEANING,
                    '0' = 1,
                    '1' = 2,
                    '2' = 3,
                    '3' = 4,
                    '4' = 5,
                    '5' = 6,
                    '6' = 7))

data_filled$SCHAPPY <- as.numeric(data_filled$SCHAPPY)
data_filled <- data_filled %>%
  mutate(SCHAPPY=recode(SCHAPPY,
                    '0' = 1,
                    '1' = 2,
                    '2' = 3,
                    '3' = 4,
                    '4' = 5,
                    '5' = 6,
                    '6' = 7))


data_filled$SCSTRESS <- as.numeric(data_filled$SCSTRESS)
data_filled <- data_filled %>%
  mutate(SCSTRESS=recode(SCSTRESS,
                    '0' = 1,
                    '1' = 2,
                    '2' = 3,
                    '3' = 4,
                    '4' = 5,
                    '5' = 6,
                    '6' = 7))
```

## Exclusions

```{r exclusions}
#include participants who answered the well-being module
data_short <- data_filled %>%
  filter(WBELIG==1) %>%
  subset(select = -c(WBELIG, RECTYPE))

#randomly selected to report well-being during leisurely activites (ATUS activity coding category 12 and 13)
data_short <- data_short %>%
  filter(ACTIVITY < 140000) %>%
  filter(ACTIVITY >=120000) %>%
  filter(SCHAPPY %in% (1:7)) %>%
  filter(SCSTRESS %in% (1:7)) %>%
  filter(MEANING %in% (1:7)) %>%
  filter(PAIDHOUR != 99) %>% #don't have information on hourly payment status 
  filter(UHRSWORKT != 9999) %>% #don't have number of hours worked
  filter(UHRSWORKT != 9995) %>%#responded as hours varied 
  filter(EARNWEEK != 0)

data_short <- data_short %>%
  group_by(CASEID) %>%
  filter(n()>=2) %>%
  mutate (CASEID=cur_group_id())%>%
  ungroup()

```

``` {r centering}
data_short <- data_short %>%
  group_by(CASEID) %>%
  mutate (DURATION_CM = mean(DURATION),
          DURATION_CMC = DURATION - DURATION_CM) %>%
  ungroup()

data_short %>%
  select(CASEID, DURATION, DURATION_CM, DURATION_CMC)


```

## Variable Description

- CASEID: unique identifying number for participant
- DAY: whether the day of the week of the diary day was a weekday (DAY = 1) or a weekend (DAY = 2)
- AGE: person's age in years as of his/her last birthday
- SEX: whether the individual is male (SEX = 1) or female (SEX = 2)
- MARST: person's marital status, either not married (MARST = 1) or married (MARST = 2)
- FULLPART: indicates whether the individual usually works full time (FULLPART = 1) or part time (FULLPART = 2).
- UHRSWORKT: total number of hours the respondent usually works per week 
- EARNWEEK: income per week 
- PAIDHOUR: whether the respondent is paid on an hourly basis (PAIDHOUD = 2) or not (PAIDHOUR = 1)
- HH_NUMOWNKIDS: the number of the respondent's own children under the age of 18
who live in the household
- DURATION: length of the activity in minutes
- SCHAPPY: seven point scale that indicates how much happiness the respondent felt during (1 - Not At all, 7 - Very)
the activity.
- SCSTRESS: seven point scale that indicates how much stress the respondent felt during
the activity (1 - Not At all, 7 - Very)
- MEANING: seven point scale that indicates how meaningful the activity was to the
respondent (1 - Not At all, 7 - Very)

## Model Equation

Lv-1:

$$\text{SCHAPPY}_{ij} = \beta_{0j} + \beta_{1j} \text{DURATION_CMC}_{ij} + e_{ij}$$

Lv-2:

$$
\begin{aligned}
  \beta_{0j} & = \gamma_{00} + \gamma_{01} \text{DURATION_CM}_j + \gamma_{02} \text{PAIDHOUR}_j + X_{3j} \gamma_{03} + ... + X_{8j} \gamma_{08} + u_{0j}  \\
  \beta_{1j} & = \gamma_{10} + \gamma_{11}\text{PAIDHOUR}_j  + u_{1j}
\end{aligned}
$$

## Descriptive Statistics

```{r des_stats}
summary(data_short)

data_short_d <- data_short %>%
  subset(select = -c(CASEID, FULLPART, WB_RESP, Leisure_All, ACTLINE, ACTIVITY, EARNWEEK, SCSTRESS, MEANING, DURATION_CM, DURATION_CMC))

apa.cor.table(data_short_d, filename="cortable2.doc", table.number=1)

data_short %>%
    select(PAIDHOUR, EARNWEEK, UHRSWORKT, DURATION) %>%
    psych::pairs.panels(ellipses = FALSE, cex = 0.2, cex.cor = 1)

data_short %>%
    select(SCHAPPY, MEANING, SCSTRESS) %>%
    psych::pairs.panels(ellipses = FALSE, cex = 0.2, cex.cor = 1)

```

## ICC

```{r ICC}
m0_hap <- lmer(SCHAPPY ~ (1 | CASEID), data = data_short)
performance::icc(m0_hap)


```
Yes, there is evidence that happiness derived from leisure varies across individuals, which accounts for 
##Propensity Score Analysis
```{r propensity-estimation}
data_short$PAIDHOUR <- as.numeric(data_short$PAIDHOUR)
data_ps <- data_short %>%
  mutate(PAIDHOUR=recode(PAIDHOUR, 
                    '1' = 0,
                    '2' = 1))

m_ps <- glmer(PAIDHOUR ~ SEX + AGE + MARST + DAY + UHRSWORKT + HH_NUMOWNKIDS + DURATION +
                (1 | CASEID), family = "binomial",
              glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 100000)),
              data = data_ps)
ps_mlm <- predict(m_ps, re.form = NULL, type = "response")

#Plot the propensity scores across the two conditions of `PAIDHOUR`
plot(factor(data_ps$PAIDHOUR), ps_mlm)

#Stratification
ps_strata <- ntile(ps_mlm, 5)
mean(data_ps$SCHAPPY[data_ps$PAIDHOUR == 1 & ps_strata == 3])
mean(data_ps$SCHAPPY[data_ps$PAIDHOUR == 0 & ps_strata == 3])
```

## Multilevel Analysis 
```{r analysis}
#random intercept
m0 <- lmer(SCHAPPY ~ PAIDHOUR + (1|CASEID), data = data_short)
summary(m0)

m02 <- lmer(SCHAPPY ~ DURATION + (1|CASEID), data = data_short)
summary (m02)

sjPlot::plot_model(m02,
                   type = "pred", terms = "DURATION",
                   show.data = TRUE, title = "",
                   dot.size = 0.5
                   )

#covariate
m_c <- lmer(SCHAPPY ~ PAIDHOUR + SEX + AGE + MARST + DAY + UHRSWORKT + HH_NUMOWNKIDS + (1 | CASEID),
           data = data_short)
summary(m_c)

#cross level
data_short %>%
  filter (CASEID %in% sample(unique(CASEID), 16)) %>%
  ggplot (aes (x = DURATION, y = SCHAPPY)) + 
  geom_point (size=0.5) +
  geom_smooth (method = "lm") + 
  facet_wrap(~CASEID)

m_cl <- lmer (SCHAPPY ~ PAIDHOUR * DURATION_CMC + DURATION_CM + (DURATION_CMC | CASEID), data = data_short)
ranova(m_cl)
summary(m_cl)

m_cl_c <- lmer(SCHAPPY ~ PAIDHOUR * DURATION_CMC  + DURATION_CM + SEX + AGE + MARST + DAY + UHRSWORKT + HH_NUMOWNKIDS + (DURATION_CMC | CASEID),
           data = data_short)
summary(m_cl_c)



msummary(list(
    "Model 1" = m0,
    "Model 2" = m_cl,
    "Model 3" = m_cl_c
))

```

## Results
```{r data-summary}
msummary(m_cl_c,
         estimate = c("{estimate} [{conf.low}, {conf.high}]"),
         statistic = NULL,  # suppress the extra rows for SEs
         shape = effect + term ~ model,
         title = "Table 1: Model coefficients")

```


```{r visualization}

m_cl_c %>%
  augment(data=data_short) %>%
  ggplot(aes (
    x = DURATION, y = .fitted, group=factor(CASEID),
    color = factor(PAIDHOUR)
  ))+
  geom_smooth(method="lm", se=FALSE, size=0.5) +
  labs (y = "Happiness from Leisure", color = "PAIDHOUR")

# sjPlot::plot_model(mb,
#                    type = "pred", terms = "PAIDHOUR",
#                    show.data = TRUE, title = "",
#                    dot.size = 0.5
#                    )
# +
#   stat_summary(
#     data = data_short, aes (x = PAIDHOUR, y = SCHAPPY),
#     fun = mean, geom = "point",
#     col = "red",
#     shape = 17,
#     size = 3,
#     alpha = 0.7
#   )

```

If hourly workers have a greater propensity to make economic evaluation of time, they should display greater sensitivity to the opportunity cost of time and as a consequence, derive less happiness from leisurely activities. To explore people's subjective experience of leisurely activities, I used the happiness measure from the ATUS well-being module. To worker's hourly work status was dummy coded "1" for non-hourly workers and "2" for hourly workers. To account for the worker's opportunity cost of time, we used the duration spent on leisurely activities and their income as an indicator of the foregone earning from spending (wasting) time on leisure. To control for potential differences between hourly and non hourly workers, we controlled for various variables: age, marital status, number of hours worked, number of children under 18 and the day of diary. To test the hypothesis that happiness derived from engaging in leisurely activities is influenced by both hourly payment status and the opportunity cost of time, I fitted a multilevel model with hourly status, duration, and logged income on happiness to explore the three-way interaction among the three variables. 

